Appendix S1

### 2022-07-05 ManuscriptTe.Rmd
ms edits

---
title: "Appendix S1"
author: "Buckley et al."
output:
  html_document: default
  pdf_document: default
  word_document: default
  rmarkdown::html_vignette: default
---

@@

```{r, include=FALSE}
C1all=read.csv("./data/microclimate_C1.csv", sep=",", na.strings = c("-49.99","NA"))
```

```{r, include=FALSE}
#ANALYZE SURFACE ROUGHNESS AND PROFILES
C1all$day=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$mday
C1all$hour=as.POSIXlt(as.character(C1all$datetime), tz='MST', format="%m/%d/%Y %H:%M")$hour

C1.dayhr<- aggregate(C1all[,c(2:8,13:22)], by=list(C1all$day, C1all$hour), FUN=mean, na.rm=TRUE)
names(C1.dayhr)[1:2]=c("day","hour")
C1.hr<- aggregate(C1all[,c(2:8,13:22)], by=list(C1all$hour), FUN=mean, na.rm=TRUE)
names(C1.hr)[1]="hour"

C1.hr= as.matrix(C1.hr)
```

@@

```{r}
z0= apply(C1.hr[, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], FUN=surface_roughness, MARGIN=1, zr=c(0.25,0.5,1.0,1.5))
```

@@

```{r}
#subset to day and hour
C1.sub= subset(C1.dayhr, C1.dayhr$day==10)
hr=10

#vector of heights
z.seq <- seq(0, 2, by = 0.1)

#temperature profiles
t.seq1= sapply(z.seq, FUN=air_temp_profile_neutral, T_r=C1.sub[hr,"Air0.5m"], z0=z0[hr], zr=0.5, T_s=C1.sub[hr,"Soil"])

t.seq2= sapply(z.seq, FUN=air_temp_profile, T_r=C1.sub[hr,"Air0.5m"], u_r=C1.hr[hr,"Wind0.5m"], zr=0.5, z0=z0[hr], T_s=C1.sub[hr,"Soil"])

t.seq3= sapply(z.seq, FUN=air_temp_profile_segment, T_r=C1.sub[hr, c("Air0.25m","Air0.5m","Air1.0m","Air1.5m")],u_r=C1.sub[hr, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], zr=c(0.25,0.5,1,1.5), z0=rep(z0[hr],4), T_s=C1.sub[hr,"Soil"])

#wind speed profiles
u.seq1= sapply(z.seq, FUN=wind_speed_profile_neutral, u_r=C1.sub[hr,"Wind0.5m"], zr=0.5, z0=z0[hr])

u.seq2= sapply(z.seq, FUN=wind_speed_profile_segment, u_r=C1.sub[hr, c("Wind0.5m","Wind1m","Wind1.5m")], zr=c(0.5,1,1.5), z0=rep(z0[hr],3))
```

@@

```{r, echo=FALSE}
#replace infinity with N
t.seq1[which(!is.finite(t.seq1))]=NA

plot(t.seq1, z.seq, type = "l", xlab = "Temperature (C)", ylab = "z (height above ground in m)")
points(t.seq2, z.seq, type = "l", lty="dashed")
points(t.seq3, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Air0.25m","Air0.5m","Air1.0m","Air1.5m")], c(0.25,0.5,1,1.5), type = "p", col="red")

plot(u.seq1, z.seq, type = "l", xlab = "windspeed (m/s)", ylab = "z (height above ground in m)")
points(u.seq2, z.seq, type = "l", lty="dotted")
points(C1.sub[hr, c("Wind0.25m","Wind0.5m","Wind1m","Wind1.5m")], c(0.25,0.5,1,1.5), type = "p", col="red")
```

@@

# Biophysical models
```{r, include=FALSE}
#Grasshopper data
spec.dat=read.csv("./data/SpecData.csv")

#read weather data
dat=read.csv("./data/DatAll_Pace2011.csv", sep=",", na.strings = c("-49.99","NA"))
#match site data
sites=read.csv("./data/SiteElevations.csv", sep=",")

#combine soil temperatures
dat$SoilTemp= rowMeans(dat[,c("SoilTemp","SoilTemperature")], na.rm = TRUE)

#omit problematic temperatures
dat$Temp[dat$Temp< -10]=NA

#add elevation
match1= match(dat$site, sites$Site)
dat$Elev=sites[match1,"Elevation"]

#calculate day of year
date.ct <- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST")
dat$J= round(as.numeric(julian(date.ct, origin = as.Date("2011-01-01") )))
dat$hour= date.ct$hour

#match site data
match1= match(dat$site, sites$Site) 
dat$lon= sites$Lon[match1]
dat$lat= sites$Lat[match1]

#Subset to August
dat.dt<- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST") #as.Date
dataug= subset(dat, dat.dt$mon==7)
#subset to day
dat= subset(dataug, dataug$Rad>5)
```

@@

dat$psi= zenith_angle(doy=dat$J,lat=dat$lat, lon=dat$lon, hour=dat$hour)

@@

We scale the air temperature ($T_r$, $^\circ$C) and wind speed ($u_r$, m/s) from a reference height of $z_r$=0.5m to grasshopper height ($z$=0.001m). The functions also require surface temperature ($T_s$, $^\circ$C), and surface roughness ($z0$, m).

```{r}
z<-0.001 #specify distance from ground

#scale temperature 
dat$Tgrass <- air_temp_profile(T_r=dat$Temp, u_r=dat$Wind, zr=0.5, z0=0.02, z=z, T_s=dat$SoilTemp)

#scale wind speed
dat$Ugrass= wind_speed_profile_neutral(u_r=dat$Wind, zr=0.5, z0=0.02,z=z)
```

```{r, include=FALSE}
#cut NAs
dat= dat[,c("Rad","Temp","Wind","SoilTemp","SanguinipesTemp","Elev","J", "hour","psi","Tgrass","Ugrass")]
dat= na.omit(dat)
```

@@

dat$Te<-Tb_grasshopper(T_a=dat$Tgrass, T_g=dat$SoilTemp, u=dat$Ugrass, H=dat$Rad, K_t=1, psi=dat$psi, l=0.0211, Acondfact=0.0, abs=0.9, r_g=0.5)

@@

```{r, include=FALSE}
#fix infinite
dat$Te[is.infinite(dat$Te)]<-NA
#ditch big numbers for now
dat$Te[dat$Te>100]<-NA
```

@@

```{r}
dat$H_L=heat_transfer_coefficient(V=dat$Ugrass,D=0.0211,K=0.025, nu= 15.3 * 10^(-6) , taxon="cylinder")

H_L1= heat_transfer_coefficient_approximation(V=dat$Ugrass,D=0.0211,K= 0.025, nu= 15.3 * 10^(-6), taxon="flyinginsect")

H_L2= heat_transfer_coefficient_simple(V=dat$Ugrass,D=0.0211, type="Gates")

@@

```{r echo=T, results = 'hide', warning=FALSE}
TeGates= rep(NA, nrow(dat) )

A= surface_area_from_length(l=0.0211) #surface area from length(m)

for(ind in 1:nrow(dat) ){

  out= try(Tb_Gates(A=A, D=0.0211/3, psa_dir=0.5, psa_ref=1-0.5, psa_air=1, psa_g=0.0, T_g=dat$SoilTemp[ind]+273, T_a=dat$Tgrass[ind]+273, Qabs=A*0.66*dat$Rad[ind], epsilon=0.95, H_L=dat$H_L[ind], ef=1.3, K=0.15), silent=TRUE)
  if(is.numeric(out))TeGates[ind]=out
}

@@

```{r, echo=FALSE, warning=FALSE}
dat$TeGates=TeGates-273.15

#drop NAs
dat= dat[which(!is.na(dat$TeGates)),]

#aggregate by hour
dat.h= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour), FUN=mean, na.rm=TRUE))
#aggregate by day and hour
dat.dh= suppressWarnings(aggregate(dat, by=list(dat$Elev, dat$hour, dat$J), FUN=mean, na.rm=TRUE))

#convert to long format
dat.s= melt(dat.h, id = c("Elev", "J","hour","psi"))

#plot
dat.pl= subset(dat.s, dat.s$variable %in% c("Temp","SoilTemp","SanguinipesTemp","Te","Tgrass","TeGates") )
dat.pl$variable= as.character(dat.pl$variable)

#change names
dat.pl[which(dat.pl$variable=="Temp"),"variable"]<-"Ta"
dat.pl[which(dat.pl$variable=="SoilTemp"),"variable"]<-"Ts"
dat.pl[which(dat.pl$variable=="SanguinipesTemp"),"variable"]<-"Te observed"
dat.pl[which(dat.pl$variable=="Te"),"variable"]<-"Te"
dat.pl[which(dat.pl$variable=="Tgrass"),"variable"]<-"Ta grasshopper height"
dat.pl[which(dat.pl$variable=="TeGates"),"variable"]<-"Te Gates"

#make factor
dat.pl$variable= factor(dat.pl$variable, levels=c("Ta","Ts","Ta grasshopper height","Te","Te Gates","Te observed"))
#make label
dat.pl$elev.lab= paste(dat.pl$Elev,"m", sep=" ")
#make Ta, Te classification
dat.pl$Ttype="input"
dat.pl$Ttype[which(dat.pl$variable %in% c("Te","Te Gates","Te observed") )]="estimates"

ggplot(data=dat.pl, aes(x=hour, y = value, color=variable))+geom_line(aes(lty=variable) )+facet_wrap(~elev.lab)+theme_classic()+ theme(legend.position="bottom")+ylim(15,65)

@@

```{r, include=FALSE}
#plot by date time
dat.dh$Jhr= dat.dh$J +dat.dh$hour/24

#convert to long format
dat.s= melt(dat.dh, id = c("Elev", "J","hour","psi","Jhr"))

#plot
dat.s= subset(dat.s, dat.s$variable %in% c("Temp","SoilTemp","SanguinipesTemp","Te","Tgrass") )
#,"TeGates"

plot.t= ggplot(data=dat.s, aes(x=Jhr, y = value, color=variable))+geom_line(aes(lty=variable))+facet_grid(Elev~1)
```

>>>>

---
title: "Appendix S1"
author: "Buckley et al."
output:
  html_document: default
  word_document: default
  pdf_document: 
    pandoc_args: --listings
    includes:
      in_header: preamble.tex
---

@@

```{r, include=FALSE}
C1all <- read.csv("./data/microclimate_C1.csv", sep = ",", na.strings = c("-49.99","NA"))
```

```{r, include=FALSE}
# ANALYZE SURFACE ROUGHNESS AND PROFILES
C1all$day <- as.POSIXlt(as.character(C1all$datetime), tz = 'MST', format = "%m/%d/%Y %H:%M")$mday
C1all$hour <- as.POSIXlt(as.character(C1all$datetime), tz = 'MST', format = "%m/%d/%Y %H:%M")$hour

C1.dayhr <- aggregate(C1all[,c(2:8,13:22)], by = list(C1all$day, C1all$hour), FUN = mean, na.rm = TRUE)
names(C1.dayhr)[1:2] <- c("day","hour")
C1.hr <- aggregate(C1all[,c(2:8,13:22)], by = list(C1all$hour), FUN = mean, na.rm = TRUE)
names(C1.hr)[1] <- "hour"

C1.hr <- as.matrix(C1.hr)
```

@@

```{r}
z0 <- apply(C1.hr[, c("Wind0.25m", "Wind0.5m", "Wind1m", "Wind1.5m")], FUN = surface_roughness, MARGIN = 1, zr = c(0.25, 0.5, 1.0, 1.5))
```

@@

```{r}
# subset to day and hour
C1.sub <- subset(C1.dayhr, C1.dayhr$day == 10)
hr <- 10

# vector of heights
z.seq <- seq(0, 2, by = 0.1)

# temperature profiles
t.seq1 <- sapply(z.seq, FUN = air_temp_profile_neutral, T_r = C1.sub[hr, "Air0.5m"], z0 = z0[hr], zr = 0.5, T_s = C1.sub[hr, "Soil"])

t.seq2 <- sapply(z.seq, FUN = air_temp_profile, T_r = C1.sub[hr, "Air0.5m"], u_r = C1.hr[hr, "Wind0.5m"], zr = 0.5, z0 = z0[hr], T_s = C1.sub[hr, "Soil"])

t.seq3 <- sapply(z.seq, FUN = air_temp_profile_segment, T_r = C1.sub[hr, c("Air0.25m", "Air0.5m", "Air1.0m", "Air1.5m")], u_r = C1.sub[hr,  c("Wind0.25m", "Wind0.5m", "Wind1m", "Wind1.5m")], zr=c(0.25, 0.5, 1, 1.5), z0 = rep(z0[hr], 4), T_s = C1.sub[hr, "Soil"])

# wind speed profiles
u.seq1 <- sapply(z.seq, FUN = wind_speed_profile_neutral, u_r = C1.sub[hr, "Wind0.5m"], zr = 0.5, z0 = z0[hr])

u.seq2 <- sapply(z.seq, FUN = wind_speed_profile_segment, u_r = C1.sub[hr, c("Wind0.5m", "Wind1m", "Wind1.5m")], zr = c(0.5, 1, 1.5), z0 = rep(z0[hr], 3))
```

@@

```{r, echo=FALSE, fig.height=4, fig.align = 'center', out.width = "75%"}
# replace infinity with N
t.seq1[which(!is.finite(t.seq1))] <- NA

plot(t.seq1, z.seq, type = "l", xlab = "Temperature (C)", ylab = "z (height above ground in m)")
points(t.seq2, z.seq, type = "l", lty = "dashed")
points(t.seq3, z.seq, type = "l", lty = "dotted")
points(C1.sub[hr, c("Air0.25m", "Air0.5m", "Air1.0m", "Air1.5m")], c(0.25, 0.5, 1, 1.5), type = "p", col = "red")

plot(u.seq1, z.seq, type = "l", xlab = "windspeed (m/s)", ylab = "z (height above ground in m)")
points(u.seq2, z.seq, type = "l", lty = "dotted")
points(C1.sub[hr, c("Wind0.25m", "Wind0.5m", "Wind1m", "Wind1.5m")], c(0.25, 0.5, 1, 1.5), type = "p", col = "red")
```

@@

# Biophysical models
```{r, include=FALSE}
#Grasshopper data
spec.dat <- read.csv("./data/SpecData.csv")

# read weather data
dat <- read.csv("./data/DatAll_Pace2011.csv", sep = ",", na.strings = c("-49.99", "NA"))
#match site data
sites <- read.csv("./data/SiteElevations.csv", sep= ",")

# combine soil temperatures
dat$SoilTemp <- rowMeans(dat[,c("SoilTemp", "SoilTemperature")], na.rm = TRUE)

# omit problematic temperatures
dat$Temp[dat$Temp< -10] <- NA

# add elevation
match1 <- match(dat$site, sites$Site)
dat$Elev <- sites[match1, "Elevation"]

# calculate day of year
date.ct <- strptime(dat$datetime, format = "%m/%d/%Y %H:%M", tz = "MST")
dat$J <- round(as.numeric(julian(date.ct, origin = as.Date("2011-01-01"))))
dat$hour <- date.ct$hour

# match site data
match1 <- match(dat$site, sites$Site) 
dat$lon <- sites$Lon[match1]
dat$lat <- sites$Lat[match1]

# Subset to August
dat.dt<- strptime(dat$datetime, format = "%m/%d/%Y %H:%M", tz = "MST") # as.Date
dataug <- subset(dat, dat.dt$mon == 7)

# subset to day
dat <- subset(dataug, dataug$Rad > 5)
```

@@

dat$psi <- zenith_angle(doy = dat$J, lat = dat$lat, lon = dat$lon, hour = dat$hour)

@@

We scale the air temperature ($T_r$, $^\circ$C) and wind speed ($u_r$, m/s) from a reference height of $z_r$ = 0.5m to grasshopper height ($z$ = 0.001m). The functions also require surface temperature ($T_s$, $^\circ$C), and surface roughness ($z0$, m).

```{r}
z <- 0.001 #specify distance from ground

# scale temperature 
dat$Tgrass <- air_temp_profile(T_r = dat$Temp, u_r = dat$Wind, zr = 0.5, z0 = 0.02, z = z, T_s = dat$SoilTemp)

# scale wind speed
dat$Ugrass <- wind_speed_profile_neutral(u_r = dat$Wind, zr = 0.5, z0 = 0.02, z = z)
```

```{r, include=FALSE}
# cut NAs
dat <- dat[,c("Rad", "Temp", "Wind", "SoilTemp", "SanguinipesTemp", "Elev", "J", "hour", "psi", "Tgrass", "Ugrass")]
dat <- na.omit(dat)
```

@@

dat$Te <- Tb_grasshopper(T_a = dat$Tgrass, T_g = dat$SoilTemp, u = dat$Ugrass, H = dat$Rad, K_t = 1, psi = dat$psi, l = 0.0211, Acondfact = 0.0, abs = 0.9, r_g = 0.5)

@@

```{r, include=FALSE}
# fix infinite
dat$Te[is.infinite(dat$Te)] <- NA

# ditch big numbers for now
dat$Te[dat$Te>100] <- NA
```

@@

```{r}
dat$H_L <- heat_transfer_coefficient(V = dat$Ugrass, D = 0.0211, K = 0.025, nu = 15.3 * 10^(-6) , taxon = "cylinder")

H_L1 <- heat_transfer_coefficient_approximation(V = dat$Ugrass, D = 0.0211, K = 0.025, nu = 15.3 * 10^(-6), taxon = "flyinginsect")

H_L2 <- heat_transfer_coefficient_simple(V = dat$Ugrass, D = 0.0211, type = "Gates")

@@

```{r, message=FALSE}
TeGates <- rep(NA, nrow(dat))

# surface area from length(m)
A <- surface_area_from_length(l = 0.0211)

for (ind in 1:nrow(dat)) {

  out <- Tb_Gates(A = A, D = 0.0211/3, psa_dir = 0.5, psa_ref = 1-0.5, psa_air = 1, psa_g = 0.0, T_g = dat$SoilTemp[ind] + 273, T_a = dat$Tgrass[ind] + 273, Qabs = A * 0.66 * dat$Rad[ind], epsilon = 0.95, H_L = dat$H_L[ind], ef = 1.3, K = 0.15)

  if (is.numeric(out)) {
    TeGates[ind] <- out
  }

}

@@

```{r, echo=FALSE, warning=FALSE}
dat$TeGates <- TeGates - 273.15

# drop NAs
dat <-  dat[which(!is.na(dat$TeGates)), ]

# aggregate by hour
dat.h <- suppressWarnings(aggregate(dat, by = list(dat$Elev, dat$hour), FUN = mean, na.rm = TRUE))

# aggregate by day and hour
dat.dh <- suppressWarnings(aggregate(dat, by = list(dat$Elev, dat$hour, dat$J), FUN = mean, na.rm = TRUE))

# convert to long format
dat.s <- melt(dat.h, id = c("Elev", "J", "hour", "psi"))

# plot
dat.pl <- subset(dat.s, dat.s$variable %in% c("Temp", "SoilTemp", "SanguinipesTemp", "Te", "Tgrass", "TeGates"))
dat.pl$variable <- as.character(dat.pl$variable)

# change names
dat.pl[which(dat.pl$variable=="Temp"),"variable"] <- "Ta"
dat.pl[which(dat.pl$variable=="SoilTemp"),"variable"] <- "Ts"
dat.pl[which(dat.pl$variable=="SanguinipesTemp"),"variable"] <- "Te observed"
dat.pl[which(dat.pl$variable=="Te"),"variable"] <- "Te"
dat.pl[which(dat.pl$variable=="Tgrass"),"variable"] <- "Ta grasshopper height"
dat.pl[which(dat.pl$variable=="TeGates"),"variable"] <- "Te Gates"

# make factor
dat.pl$variable <- factor(dat.pl$variable, levels = c("Ta", "Ts", "Ta grasshopper height", "Te", "Te Gates", "Te observed"))

# make label
dat.pl$elev.lab <- paste(dat.pl$Elev, "m", sep = " ")

# make Ta, Te classification
dat.pl$Ttype <- "input"
dat.pl$Ttype[which(dat.pl$variable %in% c("Te","Te Gates","Te observed") )] <- "estimates"

ggplot(data = dat.pl, aes(x=hour, y = value, color = variable))+ geom_line(aes(lty = variable))+ facet_wrap( ~ elev.lab) + theme_classic()+ theme(legend.position = "bottom") + ylim(15, 65)

@@

```{r, include=FALSE}
# plot by date time
dat.dh$Jhr <- dat.dh$J +dat.dh$hour/24

# convert to long format
dat.s <- melt(dat.dh, id = c("Elev", "J", "hour", "psi", "Jhr"))

# plot

dat.s <- subset(dat.s, dat.s$variable %in% c("Temp", "SoilTemp", "SanguinipesTemp", "Te", "Tgrass") )

# "TeGates"

plot.t <- ggplot(data = dat.s, aes(x = Jhr, y = value, color=variable)) + geom_line(aes(lty = variable)) + facet_grid(Elev ~ 1)
```
